### 2018-09-25 manuscript/templates4Authors/latex_template/author.tex
starting manuscript;figures

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a contributed volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[graybox]{svmult}

% choose options for [] as required from the list
% in the Reference Guide

\usepackage{mathptmx}       % selects Times Roman as basic font
\usepackage{helvet}         % selects Helvetica as sans-serif font
\usepackage{courier}        % selects Courier as typewriter font
\usepackage{type1cm}        % activate if the above 3 fonts are
                            % not available on your system
%
\usepackage{makeidx}         % allows index generation
\usepackage{graphicx}        % standard LaTeX graphics tool
                             % when including figure files
\usepackage{multicol}        % used for the two-column index
\usepackage[bottom]{footmisc}% places footnotes at page bottom

% see the list of further useful packages
% in the Reference Guide

\makeindex             % used for the subject index
                       % please use the style svind.ist with
                       % your makeindex program

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% track changes
\usepackage[normalem]{ulem}
\usepackage[dvipsnames]{xcolor}
\newcommand{\claudia}[2]{\sout{#1}\textcolor{magenta}{\textbf{#2}}}

\begin{document}

\title*{Prediction of functional markers of mass cytometry data via
  deep learning}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
%\thanks{Both authors contributed equally to this manuscript}
\author{Claudia Sol\'is-Lemus, Xin Ma, Maxwell Hostetter II, Suprateek
  Kundu, Peng Qiu, Daniel Pimentel-Alarc\'on}
%\author{Xin Ma, Claudia Sol\'is-Lemus, Maxwell Hostetter II, Suprateek
%  Kundu, Daniel Pimentel-Alarc\'on}
%\author{xxx\thanks{work the same}, xxs$^*$}% \footnotemark[1]}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{Claudia Sol\'is-Lemus \at Emory University, Atlanta, GA, \email{csolisl@emory.edu}
\and Xin Ma \at Emory University, Atlanta, GA,
\email{xin.ma@emory.edu}
\and Maxwell Hostetter II \at Georgia State University, Atlanta, GA,
\email{mhostetter1@student.gsu.edu}
\and Suprateek Kundu \at Emory University, Atlanta, GA,
\email{suprateek.kundu@emory.edu}
\and Peng Qiu \at Georgia Institute of Technology, Atlanta, GA,
\email{peng.qiu@bme.gatech.edu}
\and Daniel Pimentel-Alarc\'on \at Georgia State University, Atlanta, GA,
\email{pimentel@gsu.edu}}
%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\maketitle

\abstract{add here}

\section{Introduction}

% from Peng Qui PQ.docx: ---
Multiparametric single-cell analysis has advanced our understanding of
diverse biological and pathological processes, providing insights into
cellular differentiation, intracellular signaling cascades and
clinical immunophenotyping. Modern flow cytometers typically provide
simultaneous single-cell measurements of up to 12 fluorescent
parameters in routine cases, and analysis of up to 30 protein
parameters has been recently made commercially available. In addition,
a next-generation mass cytometry platform (CyTOF) has become
commercially available, which allows routine measurement of 50 or more
protein markers.

Despite the technological advances in acquiring an
increasing number of parameters per single cell, approaches for
analyzing such complex data lag behind. The existing approaches are
often subjective and labor-intensive. For example, the widely used
gating approach identifies cell types by user-defined sequences of
nested 2-D plots. There have been efforts to develop clustering
algorithms (eg, flowMeans, flowSOM, X-shift) and dimension reduction
algorithms (eg, SPADE, tSNE, Scaffold). However, there is still huge
space for developing new methods to ask new questions in this field.
% ---

Recently, deep learning models are revolutionizing the fields of
precision medicine, data mining, astronomy, human-computer
interactions, among many others, by becoming a major discovery force
in science due to the unprecedented accuracy in prediction.  Here, we
explore neural network models to predict functional markers (internal
phosphoproteins) with identify markers (cell surface proteins).

On one side, identify markers define a cell type (e.g., B cell, T
cell, monocytes, MSC), and on the other side, expressions of
functional markers identify the cell's activity (e.g., quiescent,
secreting cytokines, proliferating, apoptosis). Since CyTOF technology
allows for the simulteanous measurement of a large number of protein
markers, most CyTOF studies measure both identity markers and
functional markers, providing data for supervised learning tools, like
neural networks.

\begin{itemize}
\item Mass cytometry (CyTOF) is an emergent technological development
  for high-dimensional multiparameter single cell analysis.
\item CyTOF provides unprecedented multidimensional immune cell
  profiling and has recently been applied to characterizing peripheral
  blood cells, Natural Killer cells in viral infections, skin cells,
  cells in celiac disease, responding phenotypes in cancer and even
  holds the promise of examination of solid tumors (add references)
\item Among the features per cell, there two classes: surface markers
  (which identify the type of cell) and functional markers (which
  identify the function of the cell)
\item Collecting the functional markers is time-consuming and
  computationally expensive (yes?), so it is desirable to devise a
  prediction scheme that can allow us to predict the funcional markers
  from the surface markers
\item Deep learning methods have achieved outstanding performance in
  various computational tasks, such as image analysis, natural
  language processing, and pattern recognition (references)
\item Moreover, recent efforts to use deep learning approaches in
  genomics and biomedical applications show their flexibility for
  handling complex problems (Cires ̧an et al., 2013; Cruz-Roa et al.,
  2014; Denas and Taylor, 2013; Fakoor et al., 2013; Leung et al.,
  2014).
\item in each FCM and CyTOF run we typically collect approximately $10^5$
  to $10^6$ cells, so that the number of instances (cells) is several
  orders of magnitude larger than the number of variables (up to 50
  markers). Therefore, developing deep learning approaches to analyze
  cytometry data is very promising.
\item We present DeepCyTOF, an integrated deep learning domain
  adaptation framework, which employs one manually gated reference
  sample and utilizes it for automated gating of the remaining samples
  in a study.
\item We use this method in the dataset bla
\item We compare the NN approach to other usual prediction methods
  like regression-based and xxx (Suprateek is checking this), and show
  how the NN approach improves the prediction by the incorporation of
  non-linear and interaction effects.
\item We show the improvement driven by NN when the effects are
  nonlinear as a powerful alternative to the usual regression techniques.
\end{itemize}

\section{Data}

In a dataset we previously published
[Bendall et al 2011; Qiu et al 2011], we measured 13 identity
(surface) markers and 18 functional markers simultaneously.  This
measurement was performed on 5 health donor bone marrow samples
untreated, and also on 24 health bone marrow samples treated by 24
different drugs.  For the moment, let’s focus on the first 5 samples.

Given the data from 1 of the 5 samples, we have observations (data
points) for roughly 250,000 cells. This 250,000*(13+18) data can be
used to build a machine learning model to predict the 18 function
markers based on the 13 identity markers. In this dataset, we roughly
know that the 13 identity markers can be used to define ~25 cell
types, and we have manually created cell type label for each cell. One
machine learning question can be predicting the cell type label based
on the 13 surface markers, however, this is not the question we are
asking here. The questions we want to ask are:

1. If we use 1 sample
to train a deep learning model which use identity markers to predict
functional markers, can this model be used (generalized) to predict
functional markers in the other 4 untreated samples?

2. Based on the
trained deep learning model, can we try to identify cell clusters? And
will our cell clusters agree with the well-accepted cell types in the
literature.

3. If we focus cells belonging to the same known cell
type, and examine the distribution of functional markers and the
correlation with the subtle variations of the identity markers among
cells of this type, are we able to see hints that this cell type may
be further divided into subtypes?

The data is highly complex, violating the linearity assumptions of
linear regression, see plot, even after the standard transformation
for CyTOF data (arcsinh with cofactor=5). \ref{exploratory}

\begin{figure}
\includegraphics[scale=0.3]{figures/surface-hist.pdf}
\includegraphics[scale=0.3]{figures/functional-hist.pdf}
\includegraphics[scale=0.3]{figures/scatterplot.pdf}
\caption{Exploratory plots of surface and functional markers. The
  histograms show a biased pattern, and the scatterplot shows non
  linearity, crucial assumptions in regression.}
\label{exploratory}
\end{figure}

\section{Materials and Methods}

We fit a neural network model to predict functional markers from
surface markers, and compare its performance to three classical
statistical methods like linear regression (unpenalized and
penalized), and random trees. Due to computational time constraints,
we could not fit a support vector regression in the dataset. We
compared the performance of the four approaches by computing the mean
square error of the predicted responses.

To fit the models, we divided the data (1,223,228 cells in 18
functional markers (responses) and 15 surface markers (predictors))
into three subsets: 250,000 rows as training data, 250,000 rows as
validation data and 223,228 rows as testing data. The training data
was used to fit each of the four models. The validation data was used
to determine the best setup (tuning parameters) of each model in terms
of mse, and the test data was used to compare the mse across methods.

We used the vector mse (put formula), and the 18 individual mse's (put
formula) on the test data to compare across methods.

\begin{equation}
MSE_{vec} = \frac{1}{n} \sum_{i=1}^n ||\hat{Y}_i - Y_i ||^2_2
\label{vectorMSEformula}
\end{equation}

where $\hat{Y}_i \in \mathbf{R}^{18}$ is the predicted vector of
responses for individual $i$, and $Y_i \in \mathbf{R}^{18}$ is the
observed vector of responses for individual $i$.

The individual MSE was then defined for predictor $k$ as
\begin{equation}
MSE_{(k)} = \frac{1}{n} \sum_{i=1}^n (\hat{Y}_{k,i} - Y_{k,i} )^2
\label{MSEformula}
\end{equation}
where $\hat{Y}_{k,i} \in \mathbf{R}$ is the $k^{th}$ predicted
response ($k=1,\cdots,18$) for individual $i$, and $Y_{k,i} \in
\mathbf{R}$ is the $k^{th}$ observed response for individual $i$.

% https://github.com/pluskid/Mocha.jl/blob/master/docs/user-guide/solver.rst
\vspace{0.25cm}
\noindent \textbf{Neural network model} We tested different network
architectures, activation functions, regularization coefficients,
solver methods, momentum policies, and learning rates. The best
network has four layers (see figure xxx) with 90, 90, 45 and 45
nodes. The network uses hyperbolic tangent as activation function,
regularization coefficient 0.0001, momentum policy at 0.8, learning
rates of (0.05, 0.01, 0.0001, 0.75).

\vspace{0.25cm}
\noindent \textbf{Linear regression (unpenalized/penalized)} We fit standard
linear regression, as well as the penalized version with LASSO penalty
under different penalization parameters. We used the
\textit{ScikitLearn} julia wrapper, with default settings.

\vspace{0.25cm}
\noindent \textbf{Decision tree and random trees} We fit one decision
regression tree with \textit{ScikitLearn} julia wrapper, with default
settings. We compared the performance of the ``mse'' criterion and the
Friedman's improvement score, deciding on the former (``mse'') which
is the default setting. We did not constraint the maximum depth of the
tree, and used as 2 the minimum number of samples required to split an
internal node. We did not constraint the maximum number of features to
consider when looking for the best split. Later, we fit 20 trees into
a random forest method, which showed good performance but not a lot of
time waiting.

\section{Results}

In Figure \ref{vectorMSE}, we have the comparison of the individual
MSE per method, per response (18 functional markers). We can see in
this plot that the neural network model outperforms the other standard
approaches. In Figure \ref{vectorMSE}, we also see that the time is
better for the neural network approach.

\begin{figure}
\centering
\includegraphics[scale=0.35]{figures/vector-mse.pdf}
\includegraphics[scale=0.35]{figures/time.pdf}
\caption{Vector MSE for all 18 predictors as in formula
  \ref{vectorMSEformula}, and running time in seconds.. We see that
  the performance of neural network is better than all the
  others, without being much slower.}
\label{vectorMSE}
\end{figure}

In Figure \ref{MSE}, we have the comparison of the joint MSE for all 18
responses. Here we can also see that the performance of neural network
is better than other approaches.

\begin{figure}
\centering
\includegraphics[scale=0.45]{figures/mse.pdf}
\caption{MSE for each 18 predictors as in formula
  \ref{MSEformula}. We see that the performance of neural
  network is better than all the others. Lines drawn only to
  appreciate the order among methods better.}
\label{MSE}
\end{figure}

\section{Discussion}

We showed that neural network is better than other approaches, mostly
due to the lack of simplifing assumptions like lienarity. The
performance is better in terms of prediction accuracy (measured by
mse) and computational time.

unlike regression, neural network takes advatange of the correlation
among responses by fitting a network for the whole response vector.

For future work, So, we have N experiments: N+1 matrices 100k by 50: B
(baseline), $E_1,...,E_N$. We want to use the baseline information to
predict the funcional markers at different experiments (see below for
more details) (also, surface markers do not change). That is, use B to
predict $E_i$ with a neural network. Using the individual 1 as
training, we can then only get baseline information for future
individuals and predict their functional markers. This is great
because this experiments are expensive, so our method would be widely
used to save money.

%\input{references} %%originally referenc

\end{document}